---
title: "Dataset selection"
author: "Jacques Marc-Antoine"
date: "January 14, 2019"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Flights dataset

### The dataset
Available as an R package "nycflights13".

Records of flights with departure time, destination, delay, model of the plane, company...
```{r cars}
library(nycflights13)
nycflights13::flights
```

The package also contain some extra data on airports (location), planes (size, engines), hourly weather in the different airports.
```{r eval=F}
nycflights13::airports
nycflights13::airlines
nycflights13::planes
nycflights13::weather
```

### Potential tasks

* Regression: Predict delay according to meteo.

* Classification: Which company are reliable? Which are not? 


## TMDB 5000 Movie Dataset

### The dataset
Available on kaggle at: https://www.kaggle.com/tmdb/tmdb-movie-metadata

Comprises 2 datasets, one for movies (ratings, budget...), one for credits (actors and crew)

```{r}
library(data.table)
movies <- fread("data/tmdb_5000_movies.csv", nThread = 4)
str(movies)
#head(movies, 3)

credits <- fread("data/tmdb_5000_credits.csv", nThread = 4)
str(credits)
#head(credits, 3)
```

These tables can be merged using titles of the movies.

### Possible tasks

* Regression: Predict popularity based on actors, budget...

* Classification: Is Nicolas Cage playing in this movie?

* NLP: Mine the short abstracts for more features?

### Related dataset

Can be crossed with: https://www.kaggle.com/rounakbanik/the-movies-dataset#movies_metadata.csv

## 120 years of Olympics

### The dataset
Available on kaggle at:https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results

It's a pretty simple dataset containing physical attributes of athletes, their sport and medals:
```{r}
athletes <- fread("data/athlete_events.csv")
head(athletes, 5)
```

### Possible tasks:

* Which physical attributes increases chance of medals in a given sport?


## New York Stock Exchange

### The dataset
Available on kaggle: https://www.kaggle.com/dgawlik/nyse#securities.csv

Info on companies (value, activity...) and time-series data (stock exchange).
```{r}
stock <- fread("data/prices.csv")
stock
securities <- fread("data/securities.csv")
head(securities,5)
```

### Possible tasks

* Time-series forecasting

