---
title: "Time Series in R - BernR Meetup"
author: "Isabel Martinez"
date: "9/9/2019"
output:
  beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, results='hide', evaluate = FALSE, message=FALSE, warning=FALSE}
# library(tsbox)
# library(ggplot2)
# library(tidyverse)
```
## Overview

Time series as data type in base R

\bigskip

Seasonal adjustment made easy

\bigskip

Visualizing data with ts.plot

\bigskip

tsbox


## Time series

Many data have a time dimension which naturally structures the data. In econ, typical examples include:

- annual profits of a company
- quarterly sales
- monhtly imports and exports
- daily stock prices or exchange rates



## Example: Swiss Stock Exchange Index SMI

```{r EuStockMarkets, echo = FALSE}
ts.plot(EuStockMarkets[, "SMI"], 
         ylab = "", xlab = "", lwd = 2, col = "red")
grid(NULL, NULL, lwd = 1, col = "gray61")
```

<!-- I did't manage to make 2-column slides, but the following code would make 16:9 slides: -->
<!-- output: -->
<!--   beamer_presentation: -->
<!-- classoption: "aspectratio=169" -->

## Time series objects in R
- Can define data as time series; very handy, e.g., to:
     - compute growth rates
     - work with lagged values
     - seasonally adjust the data
- Base R knows the object class ts
- Just have to tell R the start date and the frequency of our data 

\pause
\medskip

- Many popular packages for time series exist, \newline
e.g., zoo and xts
- They all come with their own syntayx...
- ... and object classes
- Great source for confusion if you're just starting out!



## Defining a ts-object: annual series
```{r, echo=TRUE}

values <- round(runif(24, min = 1, max = 50))

# annual series
ts(values, frequency = 1, start = 1959)

```

## Defining a ts-object: quarterly series
- Definition of frequency: observations per year

- Quarterly data: 4 observations per year $\rightarrow$ frequency = 4

```{r, echo=TRUE}

# quarterly series
ts(values, frequency = 4, start = c(1959, 2))

```



## Defining a ts-object: monthly series

```{r, echo=TRUE}

# monthly series
ts(values, frequency = 12, start = c(1959, 1))

```


## Defining a ts-object: decennial series

```{r, echo=TRUE}
values <- round(runif(12, min = 1, max = 50))

# decennial series
ts(values, frequency = 0.1, start = 1860)


```


## Useful functions with ts objects

```{r , echo=TRUE, results="hide"}
# what is the time frame of a given ts?
time(AirPassengers)

# what is the frequency of a given ts?
frequency(AirPassengers)
```


```{r, echo=TRUE}
# select a certain time window
window(AirPassengers,
       start = c(1949, 10), 
       end = c(1951, 7))
```
 
```{r,  results="hide"}
# work with lags and leading values
stats::lag(AirPassengers, 12)

```



## Dealing with seasonality
- TS often have systematic seasonal patterns 

- These patterns add noise to our measurement

- We are often more interested in what the trend value of a variable is, rather than its actual value in a given month

- Seasonal patterns make it hard to compare values of a variable at two different points in time

## Example: Monthly totals of international airline passengers
```{r AirPassengers, echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}

ts.plot(AirPassengers, 
         ylab = "", xlab = "", lwd = 2, col = "black")
grid(NULL, NULL, lwd = 1, col = "gray61")

```
## Seasonal adjutment made easy with "seasonal"
- Great package by Christoph Sax for seasonal adjustment

- Method: X-13ARIMA-SEATS (US-Census Bureau)

- Objects must be of ts class

- Quick and easy, no need to set model parameters \newline
(but you can!)

```{r, echo=TRUE}
library(seasonal)
??seas

# obtain seasonally adjusted series
AirPassengers_sa <- final(seas(AirPassengers))

# obtain trend series
AirPassengers_t <- trend(seas(AirPassengers))

```


## Example: Monthly totals of international airline passengers
```{r , echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}

ts.plot(AirPassengers, AirPassengers_sa,
         ylab = "", xlab = "", lwd = 2, col = c("black", "red"))
grid(NULL, NULL, lwd = 1, col = "gray61")

```

## Example: Monthly totals of international airline passengers
```{r , echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}

ts.plot( AirPassengers_sa, AirPassengers_t,
         ylab = "", xlab = "", lwd = c(2, 1.5), col = c("red", "blue"))
grid(NULL, NULL, lwd = 1, col = "gray61")

```



## Plotting ts objects: ts.plot (base R)
- ts.plot (base R) quickly plots ts objects

```{r, echo = TRUE, fig.height = 3.5, fig.width = 4.5, fig.align = "center"}
ts.plot(AirPassengers)
```



## Modifying basic aspects of the graph
```{r, echo = TRUE, results="hide", fig.show = 'hide'}
ts.plot(AirPassengers, 
        ylab = "International airline passengers", 
        xlab = "", 
        lwd = 2, 
        col = "red")
```

## Result: Modifying basic aspects of the graph
```{r, echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
ts.plot(AirPassengers, 
        ylab = "International airline passengers",
        xlab = "", 
        lwd = 2, col = "red")
```



## Adding a grid 
```{r, echo = TRUE, results="hide", fig.show = 'hide'}
ts.plot(AirPassengers, 
        ylab = "Internat. airline passengers", 
        xlab = "", 
        lwd = 2, 
        col = "red")

grid(NULL, NULL, lwd = 1, col = "gray61")

```

## Result: Adding a grid 
```{r, echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
ts.plot(AirPassengers, 
        ylab = "Internat. airline passengers",
        xlab = "", 
        lwd = 2, col = "red")

grid(NULL, NULL, lwd = 1, col = "gray61")

```



## Multiple time series and legends 
```{r, echo = TRUE, results="hide", fig.show = 'hide'}
ts.plot(EuStockMarkets[, "SMI"], 
             EuStockMarkets[, "CAC"], 
        ylab = "Closing Price Index", 
        xlab = "", 
        lwd = 2, 
        col = c("red", "blue"))

grid(NULL, NULL, lwd = 1, col = "gray61")

legend("topleft", inset=c(0.01, 0.01), ncol = 1,
       legend = c ("SMI", "CAC"),
       col = c("red", "blue"),
       lwd = 2, bg = "white" )
```


## Result: Multiple time series and legends 
```{r, echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
ts.plot(EuStockMarkets[, "SMI"], 
             EuStockMarkets[, "CAC"], 
        ylab = "Closing Price Index", 
        xlab = "", 
        lwd = 2, 
        col = c("red", "blue"))
grid(NULL, NULL, lwd = 1, col = "gray61")
legend("topleft", inset=c(0.01, 0.01), ncol = 1,
       legend = c ("SMI", "CAC"),
       col = c("red", "blue"),
       lwd = 2, bg = "white" )

```


## Second axis
```{r, echo = TRUE, eval=FALSE, results="hide", fig.show = 'hide'}
# we overlay 2 plots
# 1st plot
ts.plot(EuStockMarkets[, "SMI"], 
        ylab = "SMI",  xlab = "", ylim = c(0, 10000),
        lwd = 2, col = "red")
# 2nd plot
par(new=T)  ##new graphic parameters
ts.plot(EuStockMarkets[, "CAC"], 
        ylab = "", xlab = "", ylim = c(0, 5000), 
        lwd = 2, col = "blue",
        gpars = list( yaxt="n"))  ##suppress 1st axis
# add 2nd axis
axis(side = 4) ##1st axis has side = 2
grid(NULL, NULL, lwd = 1, col = "gray61")
legend("topleft", inset=c(0.01, 0.01), ncol = 1,
       legend = c ("SMI", "CAC"),
       col = c("red", "blue"), lwd = 2, bg = "white" )
```


## Result: Second axis
```{r, echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
ts.plot(EuStockMarkets[, "SMI"], 
        ylab = "",  xlab = "", ylim = c(0, 10000),
        lwd = 2, col = "red")
par(new=T)
ts.plot(EuStockMarkets[, "CAC"], 
        ylab = "", xlab = "", ylim = c(0, 5000), 
        lwd = 2, col = "blue",
        gpars = list( yaxt="n"))
# add 2nd axis and same color as line
axis(side = 4) ##1st axis: side = 2
grid(NULL, NULL, lwd = 1, col = "gray61")
legend("topleft", inset=c(0.01, 0.01), ncol = 1,
       legend = c ("SMI", "CAC"),
       col = c("red", "blue"),
       lwd = 2, bg = "white" )
```




## Multiple axes with colors
```{r, echo = TRUE, results="hide", fig.show = 'hide'}
ts.plot(EuStockMarkets[, "SMI"], 
        ylab = "",  xlab = "", ylim = c(0, 10000),
        lwd = 2, col = "red")
par(new=T)
ts.plot(EuStockMarkets[, "CAC"], 
        ylab = "", xlab = "", ylim = c(0, 5000), 
        lwd = 2, col = "blue",
        gpars = list( yaxt="n"))
# add 2nd axis and use same color as data it represents
axis(side = 4, 
     col="blue", col.ticks="blue", col.axis="blue") 
grid(NULL, NULL, lwd = 1, col = "gray61")
legend("topleft", inset=c(0.01, 0.01), ncol = 1,
       legend = c ("SMI", "CAC"),
       col = c("red", "blue"), lwd = 2, bg = "white" )
```


## Result: Multiple axes with colors
```{r, echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
ts.plot(EuStockMarkets[, "SMI"], 
        ylab = "",  xlab = "", ylim = c(0, 10000),
        lwd = 2, col = "red")
par(new=T)
ts.plot(EuStockMarkets[, "CAC"], 
        ylab = "", xlab = "", ylim = c(0, 5000), 
        lwd = 2, col = "blue",
        gpars = list( yaxt="n"))
# add 2nd axis and same color as line
axis(side = 4, 
     col="blue", col.ticks="blue", col.axis="blue") 
grid(NULL, NULL, lwd = 1, col = "gray61")
legend("topleft", inset=c(0.01, 0.01), ncol = 1,
       legend = c ("SMI", "CAC"),
       col = c("red", "blue"),
       lwd = 2, bg = "white" )
```



## Time series of the world unite! The tsbox package
- Many time series packages exist, all have own object class, \newline
  e.g., zoo, xts, tsibble, timeSeries, ...
  
```{r, echo=TRUE, eval=FALSE}
library(tsbox) ## (by Christoph Sax) 
```
- provides a set of tools that are agnostic towards existing standards
- even handles time series as plain data frames, allowing for dplyr or data.table workflow
\medskip

- tsbox does many more things, especially:
    - Convert everything into everything 
    - Change frequency with *ts_frequency*
    - Create indices with *ts_index*
    - Forecasting with *ts_forecast*
    - Seasonal adjustment with *ts_seas*


## Plotting with tsbox: ts_plot and ts_ggplot

- ts.plot can only deal with ts objects
- ts.plot cannot plot different frequencies

\medskip

- ts_plot 
    - plots different frequencies
    - fast and simple
    - for all time series classes, not only ts
    - limited customizability

\medskip

- ts_ggplot 
    - same syntax and similar plots as ts_plot
    - $+$ ggplot2 graphic system
    - can be highly customized

## Example: ts_plot

```{r, echo = TRUE, eval = FALSE, fig.show=FALSE}
library(tsbox)
ts_plot( Passengers = AirPassengers,
         "Annual US revenue" = airmiles/100,
  title = "Airline Data",
  subtitle = "The classic R sample data"
)
```


## Example: ts_plot

```{r, echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
library(tsbox)
ts_plot( Passengers = AirPassengers,
         "Annual US revenue" = airmiles/100,
  title = "Airline Data",
  subtitle = "The classic R sample data"
)
```